On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-03-31 manuscripts/PNAS/chazthm_pnas.tex
more SR

\significancestatement{Many naturally occurring signals (visual scenery, speech, electrocardiograms, etc.) can be characterized as parsimonious combinations of elementary waveforms drawn from a large dictionary. Absent analytic descriptions of these signal classes, it falls on algorithms from machine learning to extract such structures given many examples from a class. We give conditions guaranteeing when the dictionary of a given size is uniquely and stably determined by data, regardless of the algorithm used to compute it. These results provide theoretical grounding for  the application of dictionary learning to recover sparse signals from unknown noisy compressive measurements.}

@@

Our main technical finding is that a sufficient diversity of measurements generated from any dictionary satisfying the spark condition 
has a stable sparse representation.  More specifically, such dictionaries are uniquely identifiable from as few as \mbox{$N = m + m(k-1){m \choose k}$} sparse linear combinations of their columns up to an error that is linear in the measurement noise (Thm.~\ref{DeterministicUniquenessTheorem}). In fact, provided $(n, m, k)$ satisfy CS inequality \eqref{CScondition}, in almost all cases the dictionary learning problem is well-posed given enough data (Cor.~\ref{ProbabilisticCor}). Importantly, these guarantees hold without assuming the recovered matrix satisfies the spark condition. Our results  also apply when only an upper bound on the number of dictionary elements is known. The explicit, algorithm-independent criteria we provide should be a useful tool in the theory of sparse dictionary learning.  % (Thm.~\ref{DeterministicUniquenessTheorem2})

@@

\begin{theorem}\label{DeterministicUniquenessTheorem}
Fix integers $n, \ell$, and $k < m$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition} and that $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} contain at least \mbox{$(k-1){\hat m \choose k}+1$} vectors supported in general linear position on each set in some $\ell$-regular $E \subseteq {[m] \choose k}$ with the SIP.  There exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$.

@@

An important consequence of Thm.~\ref{DeterministicUniquenessTheorem} is that \eqref{def1} is guaranteed provided $\varepsilon$ does not exceed: [*** TODO: verify this is still correct ***]
\begin{align}\label{epsdel}
\varepsilon(\delta_1, \delta_2) := \min \left\{ \frac{\delta_1}{ C_1 }, \frac{ \delta_2 (L_{2k}(\mathbf{A}) -  C_1\varepsilon)}{ 1 + C_1 \left( \max_{i \in [N]} \|\mathbf{x}_i\|_1  + \delta_2 \right) } \right\}.
\end{align}

@@

\begin{corollary}\label{DeterministicUniquenessCorollary}
Given $n, m$, $k < m$, and a regular hypergraph $E \subseteq {[m] \choose k}$ with the SIP, there are $N =  |E| \left[ 1 + (k-1){m \choose k} \right]$ vectors \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} such that every matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} generates a dataset $Y = \{\mathbf{A}\mathbf{x}_1, \ldots, \mathbf{A}\mathbf{x}_N\}$ with a stable $k$-sparse representation in $m$ dimensions.
\end{corollary}
\begin{proof}
Given Thm.\ref{DeterministicUniquenessTheorem}, one needs only to produce sparse codes in general linear position.  However, this is straightforward with a ``Vandermonde'' matrix construction (e.g., see \cite{Hillar15}).
\end{proof}

@@

\begin{theorem}
Fix $n, \ell$, and $k < m$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition} and that $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} have at least \mbox{$k \left[ 1 + (k-1){ m \choose k} \right]$} vectors supported in general linear position on each set in an $\ell$-regular $E \subseteq {[m] \choose k}$ with the SIP.  Given any $\delta_1, \delta_2 \geq 0$ and $\varepsilon \leq \varepsilon(\delta_1, \delta_2)$, the following holds.  % (\eqref{epsdel}).
% $\varepsilon < L_2(\mathbf{A}) / C_1$.
%There exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$.
Every solution $\mathbf{B}, \mathbf{\hat x}_i, \ldots, \mathbf{\hat x}_N$ to the constrained $\ell_0$ optimization problem:
\begin{align}\label{minsum}
\arg \min \sum_{i = 1}^N \|\mathbf{\hat x}_i\|_0 \ \ \text{s.t.} \ \ \|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon, \ \ \text{$i \in [N]$},
\end{align}
necessarily satisfies the recovery inequalities in \eqref{def1}. % \eqref{Cstable} and \eqref{b-PDa}.
\end{theorem}

\begin{proof}
We derive a lower bound on the number of $k$-sparse vectors $\mathbf{\hat x}_i$ and then apply Thm.~\ref{DeterministicUniquenessCorollary}. 
Let $n_p$ be the number of $\mathbf{\hat x}_i$ with $\|\mathbf{\hat x}_i\|_0 = p$.  %, so that $\sum_{p = 1}^{m} n_p = N$. 
Since the $\mathbf{x}_i$ are all $k$-sparse, by \eqref{minsum} we have:
\begin{align}
k \sum_{p = 1}^{m} n_p = kN \geq \sum_{i=1}^N \|\mathbf{x}_i\|_0 \geq \sum_{i=1}^N \|\mathbf{\hat x}_i\|_0 = \sum_{p=1}^{m} p n_p 
\end{align}
\[\implies \sum_{p = 1}^k (k-p)n_p \geq \sum_{p = k+1}^{\hat m} (p-k) n_p.\]
Hence,
\begin{align}
(k-1) \sum_{p =1}^k n_p \geq \sum_{p = 1}^k (k-p)n_p \geq \sum_{p = k+1}^{m} (p-k) n_p
\end{align}
\[ \geq \sum_{p = k+1}^{m} n_p = N - \sum_{p = 1}^k n_p \implies \sum_{p =1}^k n_p \geq N / k.\]

In particular, at least $N / k$ of the $\mathbf{\hat x}_i$ are $k$-sparse. Therefore, there is some $N' \subseteq [N]$ for which for each $S \in E$ there are $(k-1){m \choose k}+1$ of the vectors $\mathbf{x}_i$ with $i \in [N']$ supported on $S$ and for which $\mathbf{\hat x}_i$ is $k$-sparse for all $i \in [N']$. By Thm.~\ref{DeterministicUniquenessCorollary}, \eqref{Cstable} and \eqref{b-PDa} follow.
\end{proof}

Another straightforward application of Thm.~\ref{DeterministicUniquenessTheorem} is a probabilistic extension that uses the following often used application of random matrix theory to compressed sensing.  A random matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfies \eqref{SparkCondition} with probability one (or with ``high probability'' for discrete variables) 
provided:
\begin{align}\label{CScondition}
n \geq \gamma k\log\left(\frac{m}{k}\right),
\end{align}
in which $\gamma$ is a positive constant that depends on the particular distribution from which the entries of $\mathbf{A}$ are sampled i.i.d. (many ensembles suffice, e.g. \cite[Sec.~4]{Baraniuk08}).

@@

% It so happens that a similar statement applies to sets of vectors with a stable sparse representation. As in \cite[Sec.~IV]{Hillar15}, consider the ``symbolic'' dataset $Y = \{\mathbf{A}\mathbf{x}_1,\ldots,\mathbf{A} \mathbf{x}_N\}$ generated by indeterminate $\mathbf{A}$ and $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N$.

@@

% ======== b - PDa =========
First note that $\|\mathbf{x}\|_1 \leq \sqrt{k} \|\mathbf{x}\|_2$ for $k$-sparse $\mathbf{x} \in \mathbb{R}^m$, which by definition of $L_k(A)$ implies the following often used inequality:
\begin{align}\label{delrho}
L_k(\mathbf{A}) \leq \frac{\|\mathbf{A}\mathbf{x}\|_2}{\sqrt{k} \|\mathbf{x}\|_2} %\leq \frac{\|\mathbf{x}\|_1}{\sqrt{k} \|\mathbf{x}\|_2} \max_{i \in [m]}\|\mathbf{A}_i\|_2 
\leq  \max_{i \in [m]}\|\mathbf{A}_i\|_2.
\end{align}

Our first step is to show how dictionary recovery \eqref{Cstable} already implies sparse code recovery \eqref{b-PDa} when $\varepsilon < L_{2k}(\mathbf{A}) / C_1$. For all $2k$-sparse $\mathbf{x} \in \mathbb{R}^m$, the triangle inequality gives \mbox{$\|(\mathbf{A}-\mathbf{BPD})\mathbf{x}\|_2  \leq C_1\varepsilon \|\mathbf{x}\|_1 \leq C_1 \varepsilon \sqrt{2k}  \|\mathbf{x}\|_2$}. Thus, 
\begin{align*}
\|\mathbf{BPD}\mathbf{x}\|_2 
&\geq | \|\mathbf{A}\mathbf{x}\|_2 - \|(\mathbf{A}-\mathbf{BPD})\mathbf{x}\|_2 | \\
&\geq \sqrt{2k} (L_{2k}(\mathbf{A}) -  C_1\varepsilon) \|\mathbf{x}\|_2,
\end{align*}
%
where we drop the absolute value due to the upper bound on $\varepsilon$. Hence, $L_{2k}(\mathbf{BPD}) \geq L_{2k}(\mathbf{A}) - C_1\varepsilon  > 0$ and \eqref{b-PDa} then follows from:
\begin{align*}
\|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i \|_1
%&\leq \sqrt{2k} |\mathbf{x}_i - D^{-1}P^{\top}\mathbf{\hat x}_i\|_2 \\
&\leq \frac{\|\mathbf{BPD}(\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i)\|_2}{L_{2k}(\mathbf{BPD})} \\
&\leq \frac{\|\mathbf{B}\mathbf{\hat x}_i - \mathbf{A}\mathbf{x}_i\|_2 + \|(\mathbf{A} - \mathbf{BPD})\mathbf{x}_i\|_2}{L_{2k}(\mathbf{BPD})} \\
&\leq \frac{\varepsilon (1+C_1 \|\mathbf{x}_i\|_1)}{L_{2k}(\mathbf{BPD})}.
%&\leq \frac{\varepsilon}{\sqrt{2k}} \left( \frac{ 1+C_1 \|\mathbf{x}_i\|_1 }{ L_{2k}(\mathbf{A}) -  C_1\varepsilon } \right).
%\leq \frac{\varepsilon }{\varepsilon_0 - \varepsilon} \left( C_1^{-1}+|\mathbf{x}_i\|_1 \right).
\end{align*}

The heart of the matter is therefore \eqref{Cstable}, which we now establish using the following approach. Pigeonholing the data into its set of possible supports with respect to the matrix $\mathbf{B}$, one sees that the column space of every $\mathbf{A}_S$ with $S \in E$ must be nearby (in a sense yet to be defined; see Def.~\ref{dDef}) the column space of some $n \times k$ submatrix of $\mathbf{B}$. \eqref{Cstable} then follows from our result in combinatorial matrix analysis (Lem.~\ref{MainLemma}).

@@

Fix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} and suppose that for some $\mathbf{B} \in \mathbb{R}^{n \times \hat m}$ ($\hat m \geq m$) and $1$-sparse $\mathbf{\hat x}_i \in \mathbb{R}^{\hat m}$ we have  $\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon < L_2(\mathbf{A}) / C_1$ for all $i$. Then there are $\hat{c}_1, \ldots, \hat{c}_m \in \mathbb{R}$ and $\pi: [m] \to [\hat m]$ with:
\begin{align}\label{1D}
\|c_j\mathbf{A}_j - \hat{c}_j\mathbf{B}_{\pi(j)}\|_2 \leq \varepsilon,\ \ \   \text{for all $j$}.
\end{align} 
Note that $\hat{c}_j \neq 0$ for all $j$ since otherwise by \eqref{delrho} and definition of $C_1$ we would have $\|c_j\mathbf{A}_j\|_2 < \min_{\ell \in [m]}\|c_{\ell}\mathbf{A}_{\ell}\|_2$.

@@

\begin{lemma}\label{MainLemma}
Fix integers $n$ and $k < m$, and suppose $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition}. There exists a constant $C_2 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_2$. If for some  $\mathbf{B} \in \mathbb{R}^{n \times m}$ and regular $E \subseteq2^{[m]}$ satisfying the singleton intersection property there exists a size-preserving map $\pi: E \mapsto 2^{[m]}$ satisfying:
\begin{align}\label{GapUpperBound}
d(\text{\rm Span}\{\mathbf{A}_{S}\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \leq \varepsilon,\ \ \   \text{for all $S \in E$},
\end{align}
%
then there exists a permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ with:
\begin{align}\label{MainLemmaBPD}
\|\mathbf{A}_j - (\mathbf{BPD})_j\|_2 \leq C_2 \varepsilon, \ \ \  \text{for } j \in [m].
\end{align}
\end{lemma}

@@

The constant $C_2$ in Lem.~\ref{MainLemma} is then given by:  % \footnote{Note that $C_2 > 0$ is well-defined since $r > 0$ by definition.}
\begin{align}\label{Cdef2}
C_2(\mathbf{A}, E) := \frac{ 2^{|E|} \max_{i \in [m]} \|\mathbf{A}_i\|_2}{ \min_{F \subseteq E} r( \{ \text{\rm Span}\{\mathbf{A}_{S}\} \}_{S \in F}) }.
\end{align}
%Note that when $k=1$ the constant $C_1$ is consistent with the value in our proof of the case $k=1$.

@@

Let $\mathbf{X}$ and $\mathbf{\hat{X}}$ be the $m \times N$ matrices with columns $\mathbf{x}_i$ and $\mathbf{\hat x}_i$, respectively. It follows from the general linear position of the $\mathbf{x}_i$ and the linear independence of every $k$ columns of $\mathbf{A}$ that $L_k(\mathbf{AX}_{J(S)}) > 0$; that is, the columns of the $n \times k$ matrix $\mathbf{AX}_{J(S)}$ are linearly independent and thus form a basis for $\text{\rm Span}\{\mathbf{A}_{S}\}$. Fixing $\mathbf{y} \in \text{\rm Span}\{\mathbf{A}_{S}\}$, there then exists a unique $\mathbf{c} = (c_1, \ldots, c_k) \in \mathbb{R}^k$ such that $\mathbf{y} = \mathbf{AX}_{J(S)}\mathbf{c}$. Letting \mbox{$\mathbf{\hat{y}} = \mathbf{B\hat{X}}_{J(S)}\mathbf{c}$}, which is in $\text{\rm Span}\{\mathbf{B}_{S'}\}$, we have:
\begin{align*}
\|\mathbf{y} - \mathbf{\hat{y}}\|_2 
&= \|\sum_{i=1}^k c_i(\mathbf{AX}_{J(S)} - \mathbf{B\hat{X}}_{J(S)})_i\|_2
\leq \varepsilon \sum_{i=1}^k |c_i| \\
&\leq \varepsilon \sqrt{k}  \|\mathbf{c}\|_2 
%\leq \frac{\varepsilon \sqrt{k}}{L(\mathbf{A}X_{J(S)})} \|\mathbf{A}X_{J(S)}\mathbf{c}\|_2 
= \frac{\varepsilon}{L_k(\mathbf{AX}_{J(S)})} \|\mathbf{y}\|_2,
\end{align*}
where the last equality is by definition of the domain-restricted matrix lower bound $L_k$. It follows that for all $S \in E$,
\begin{align}\label{rhs222}
d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_{S'}\}) 
\leq \frac{\varepsilon}{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) }.
%\leq \frac{\varepsilon}{L(\mathbf{AX}_{J(S)})}.
%&< \left( \frac{ L_2(\mathbf{A}) }{ 2^{|E|} \max_{j \in [m]} \|\mathbf{A}_j\|_2 } \right) \frac{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) } { L(\mathbf{AX}_{J(S)}) } \min_{\substack{F \subseteq E}} r( \{ \text{\rm Span}\{\mathbf{A}_{S}\} \}_{S \in F}) \\
\end{align}
The result then follows by  Lem.~\ref{MainLemma}, with the map $\pi: E \to {[m] \choose k}$ defined by the association $S \mapsto S'$ above.
%\begin{align}
%\delta= \frac{\varepsilon}{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) } = \frac{ C \varepsilon }{ \tilde C } < \frac{L_2(\mathbf{A})}{\tilde C}
%\end{align}
\end{proof}

>>>>

\significancestatement{Many naturally occurring signals (visual scenery, speech, electrocardiograms, etc.) can be characterized as parsimonious combinations of elementary waveforms drawn from a large dictionary. Absent analytic descriptions of these signal classes, it falls on algorithms from machine learning to extract such structures given many examples from a class. We give conditions guaranteeing when a dictionary of a given size is uniquely and stably determined by data, regardless of the algorithm used to compute it. These results provide theoretical grounding for  the application of dictionary learning to recovering sparse signals from unknown noisy compressive measurements.}

@@

Our main technical finding is that a sufficient diversity of measurements generated from any dictionary satisfying the spark condition 
has a stable sparse representation.  More specifically, such dictionaries are uniquely identifiable from as few as \mbox{$N = m(k-1){m \choose k} + m$} sparse linear combinations of their columns up to an error that is linear in the measurement noise (Thm.~\ref{DeterministicUniquenessTheorem}). In fact, provided $(n, m, k)$ satisfy CS inequality \eqref{CScondition}, in almost all cases the dictionary learning problem is well-posed given enough data (Cor.~\ref{ProbabilisticCor}). Importantly, these guarantees hold without assuming the recovered matrix satisfies the spark condition. Our results  also apply when only an upper bound on the number of dictionary elements is known. The explicit, algorithm-independent criteria we provide should be a useful tool in the theory of sparse dictionary learning.  % (Thm.~\ref{DeterministicUniquenessTheorem2})

@@

\begin{theorem}\label{DeterministicUniquenessTheorem}
Fix integers $n, \ell$, and $k < m$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition} and that $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} contain at least \mbox{$(k-1){m \choose k}+1$} vectors supported in general linear position on each set in some $\ell$-regular $E \subseteq {[m] \choose k}$ with the SIP.  There exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$.

@@

An important consequence of Thm.~\ref{DeterministicUniquenessTheorem} is that \eqref{def1} is guaranteed provided $\varepsilon$ does not exceed: % [*** TODO: verify this is still correct ***]
\begin{align}\label{epsdel}
\varepsilon(\delta_1, \delta_2) := \min \left\{ \frac{\delta_1}{ C_1 }, \frac{ \delta_2 (L_{2k}(\mathbf{A}) -  C_1\varepsilon)}{ 1 + C_1 \left( \max_{i \in [N]} \|\mathbf{x}_i\|_1  + \delta_2 \right) } \right\}.
\end{align}

@@

\begin{corollary}\label{DeterministicUniquenessCorollary}
Given $n, m$, $k < m$, and a regular hypergraph $E \subseteq {[m] \choose k}$ with the SIP, there are $N =  |E| \left[ (k-1){m \choose k} + 1  \right]$ vectors \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} such that every matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} generates a dataset $Y = \{\mathbf{A}\mathbf{x}_1, \ldots, \mathbf{A}\mathbf{x}_N\}$ with a stable $k$-sparse representation in $m$ dimensions.
\end{corollary}
\begin{proof}
Given Thm.\ref{DeterministicUniquenessTheorem}, one needs only to produce sparse codes in general linear position.  However, this is straightforward with a ``Vandermonde'' matrix construction (e.g., see \cite{Hillar15}).
\end{proof}

@@

\begin{theorem}
Fix $n, \ell$, and $k < m$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition} and that $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} have at least \mbox{$k \left[ (k-1){ m \choose k} + 1 \right]$} vectors supported in general linear position on each set in an $\ell$-regular $E \subseteq {[m] \choose k}$ with the SIP.  Given any $\delta_1, \delta_2 \geq 0$ and $\varepsilon \leq \varepsilon(\delta_1, \delta_2)$, the following holds.  % (\eqref{epsdel}).
% $\varepsilon < L_2(\mathbf{A}) / C_1$.
%There exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$.
Every solution $\mathbf{B}, \mathbf{\hat x}_i, \ldots, \mathbf{\hat x}_N$ to the constrained $\ell_0$ optimization problem:
\begin{align}\label{minsum}
\arg \min \sum_{i = 1}^N \|\mathbf{\hat x}_i\|_0 \ \ \text{s.t.} \ \ \|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon, \ \ \text{$i \in [N]$},
\end{align}
necessarily satisfies the recovery inequalities in \eqref{def1}. % \eqref{Cstable} and \eqref{b-PDa}.
\end{theorem}

\begin{proof}
We derive a lower bound on the number of $k$-sparse  $\mathbf{\hat x}_i$ and then apply Thm.~\ref{DeterministicUniquenessCorollary}. 
[*** This proof still has a gap ***]
Let $n_p$ be the number of $\mathbf{\hat x}_i$ with $\|\mathbf{\hat x}_i\|_0 = p$.  %, so that $\sum_{p = 1}^{m} n_p = N$. 
Since the $\mathbf{x}_i$ are all $k$-sparse, by \eqref{minsum} we have:
\begin{align}
k \sum_{p = 1}^{m} n_p = kN \geq \sum_{i=1}^N \|\mathbf{x}_i\|_0 \geq \sum_{i=1}^N \|\mathbf{\hat x}_i\|_0 = \sum_{p=1}^{m} p n_p 
\end{align}
\[\implies \sum_{p = 1}^k (k-p)n_p \geq \sum_{p = k+1}^{m} (p-k) n_p.\]
Hence,
\begin{align}
(k-1) \sum_{p =1}^k n_p \geq \sum_{p = 1}^k (k-p)n_p \geq \sum_{p = k+1}^{m} (p-k) n_p
\end{align}
\[ \geq \sum_{p = k+1}^{m} n_p = N - \sum_{p = 1}^k n_p \implies \sum_{p =1}^k n_p \geq N / k.\]

In particular, at least $N / k$ of the $\mathbf{\hat x}_i$ are $k$-sparse. Therefore, there is some $N' \subseteq [N]$ for which for each $S \in E$ there are $(k-1){m \choose k}+1$ of the vectors $\mathbf{x}_i$ with $i \in [N']$ supported on $S$ and for which $\mathbf{\hat x}_i$ is $k$-sparse for all $i \in [N']$.  The result now follows directly from Thm.~\ref{DeterministicUniquenessCorollary}. % \eqref{Cstable} and \eqref{b-PDa} follow.
\end{proof}

Another straightforward application of Thm.~\ref{DeterministicUniquenessTheorem} is a probabilistic extension, which takes advantage of the following well-known application of random matrix theory to compressed sensing.  A random $n \times m$ matrix obeys \eqref{SparkCondition} with probability one (or ``high probability'' for discrete variables) 
provided:
\begin{align}\label{CScondition}
n \geq \gamma k\log\left(\frac{m}{k}\right),
\end{align}
in which $\gamma >0$ is a constant that depends on the particular distribution from which the entries of $\mathbf{A}$ are sampled i.i.d. (many ensembles suffice, e.g. see \cite[Sec.~4]{Baraniuk08}).

@@

A similar phenomenon applies to sets of vectors with a stable sparse representation. As in \cite[Sec.~IV]{Hillar15}, consider the ``symbolic'' dataset $Y = \{\mathbf{A}\mathbf{x}_1,\ldots,\mathbf{A} \mathbf{x}_N\}$ generated by indeterminate $\mathbf{A}$ and $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N$.

@@

% ======== b - PDa =========
First note that $\|\mathbf{x}\|_1 \leq \sqrt{k} \|\mathbf{x}\|_2$ for $k$-sparse $\mathbf{x} \in \mathbb{R}^m$, which by definition of $L_k(A)$ implies the following often used inequality:
\begin{align}\label{delrho}
L_k(\mathbf{A}) \leq \frac{\|\mathbf{A}\mathbf{x}\|_2}{\sqrt{k} \|\mathbf{x}\|_2} %\leq \frac{\|\mathbf{x}\|_1}{\sqrt{k} \|\mathbf{x}\|_2} \max_{i \in [m]}\|\mathbf{A}_i\|_2 
\leq  \max_{j \in [m]}\|\mathbf{A}_j\|_2.
\end{align}

Our first step is to show how dictionary recovery \eqref{Cstable} already implies sparse code recovery \eqref{b-PDa} when $\varepsilon < L_{2k}(\mathbf{A}) / C_1$. For all $2k$-sparse $\mathbf{x} \in \mathbb{R}^m$, the triangle inequality gives \mbox{$\|(\mathbf{A}-\mathbf{BPD})\mathbf{x}\|_2  \leq C_1\varepsilon \|\mathbf{x}\|_1 \leq C_1 \varepsilon \sqrt{2k}  \|\mathbf{x}\|_2$}. Thus, 
\begin{align*}
\|\mathbf{BPD}\mathbf{x}\|_2 
&\geq | \|\mathbf{A}\mathbf{x}\|_2 - \|(\mathbf{A}-\mathbf{BPD})\mathbf{x}\|_2 | \\
&\geq \sqrt{2k} (L_{2k}(\mathbf{A}) -  C_1\varepsilon) \|\mathbf{x}\|_2,
\end{align*}
%
%where we drop the absolute value due to the upper bound on $\varepsilon$. 
Hence, $L_{2k}(\mathbf{BPD}) \geq L_{2k}(\mathbf{A}) - C_1\varepsilon  > 0$ and \eqref{b-PDa} then follows from:
\begin{align*}
\|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i \|_1
%&\leq \sqrt{2k} |\mathbf{x}_i - D^{-1}P^{\top}\mathbf{\hat x}_i\|_2 \\
&\leq \frac{\|\mathbf{BPD}(\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i)\|_2}{L_{2k}(\mathbf{BPD})} \\
&\leq \frac{\|\mathbf{B}\mathbf{\hat x}_i - \mathbf{A}\mathbf{x}_i\|_2 + \|(\mathbf{A} - \mathbf{BPD})\mathbf{x}_i\|_2}{L_{2k}(\mathbf{BPD})} \\
&\leq \frac{\varepsilon (1+C_1 \|\mathbf{x}_i\|_1)}{L_{2k}(\mathbf{BPD})}.
%&\leq \frac{\varepsilon}{\sqrt{2k}} \left( \frac{ 1+C_1 \|\mathbf{x}_i\|_1 }{ L_{2k}(\mathbf{A}) -  C_1\varepsilon } \right).
%\leq \frac{\varepsilon }{\varepsilon_0 - \varepsilon} \left( C_1^{-1}+|\mathbf{x}_i\|_1 \right).
\end{align*}

The heart of the matter is therefore \eqref{Cstable}, which we now finally establish, but first in the important special case $k = 1$.

@@

Fix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} and suppose that for some $\mathbf{B} \in \mathbb{R}^{n \times \hat m}$ ($\hat m \geq m$) and $1$-sparse $\mathbf{\hat x}_i \in \mathbb{R}^{\hat m}$ we have  $\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon < L_2(\mathbf{A}) / C_1$ for all $i$. Then, there are $\hat{c}_1, \ldots, \hat{c}_m \in \mathbb{R}$ and a map $\pi: [m] \to [\hat m]$ with:
\begin{align}\label{1D}
\|c_j\mathbf{A}_j - \hat{c}_j\mathbf{B}_{\pi(j)}\|_2 \leq \varepsilon,\ \ \   \text{for all $j$}.
\end{align} 
Note that $\hat{c}_j \neq 0$ since otherwise by \eqref{delrho} and definition of $C_1$ we would have $\|c_j\mathbf{A}_j\|_2 < \min_{\ell \in [m]}\|c_{\ell}\mathbf{A}_{\ell}\|_2$.  [*** TODO:  This doesn't seem correct ***]

@@

\begin{lemma}\label{MainLemma}
Fix integers $n$ and $k < m$, and suppose $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition}. There exists a constant $C_2 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_2$. If for some  $\mathbf{B} \in \mathbb{R}^{n \times m}$ and regular $E \subseteq {[m] \choose k}$ satisfying the SIP there exists a map $\pi: E \mapsto {[m] \choose k}$ satisfying:
\begin{align}\label{GapUpperBound}
d(\text{\rm Span}\{\mathbf{A}_{S}\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \leq \varepsilon,\ \ \   \text{for all $S \in E$},
\end{align}
%
then there exists a permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ with:
\begin{align}\label{MainLemmaBPD}
\|\mathbf{A}_j - (\mathbf{BPD})_j\|_2 \leq C_2 \varepsilon, \ \ \  \text{for all } j \in [m].
\end{align}
\end{lemma}

@@

The constant $C_2$ in Lem.~\ref{MainLemma} can then be expressed as:  % \footnote{Note that $C_2 > 0$ is well-defined since $r > 0$ by definition.}
\begin{align}\label{Cdef2}
C_2(\mathbf{A}, E) := \frac{ 2^{|E|} \max_{i \in [m]} \|\mathbf{A}_i\|_2}{ \min_{F \subseteq E} r( \{ \text{\rm Span}\{\mathbf{A}_{S}\} \}_{S \in F}) }.
\end{align}
%Note that when $k=1$ the constant $C_1$ is consistent with the value in our proof of the case $k=1$.

@@

Let $\mathbf{X}$ and $\mathbf{\hat{X}}$ be the $m \times N$ matrices with columns $\mathbf{x}_i$ and $\mathbf{\hat x}_i$, respectively. It follows from the general linear position of the $\mathbf{x}_i$ and the linear independence of every $k$ columns of $\mathbf{A}$ that $L_k(\mathbf{AX}_{J(S)}) > 0$; that is, the columns of the $n \times k$ matrix $\mathbf{AX}_{J(S)}$ are linearly independent and thus form a basis for $\text{\rm Span}\{\mathbf{A}_{S}\}$. Fixing $\mathbf{y} \in \text{\rm Span}\{\mathbf{A}_{S}\}$, there then exists a unique $\mathbf{c} = (c_1, \ldots, c_k) \in \mathbb{R}^k$ such that $\mathbf{y} = \mathbf{AX}_{J(S)}\mathbf{c}$. Letting \mbox{$\mathbf{\hat{y}} = \mathbf{B\hat{X}}_{J(S)}\mathbf{c}$}, which is in $\text{\rm Span}\{\mathbf{B}_{S'}\}$, we have:
\begin{align*}
\|\mathbf{y} - \mathbf{\hat{y}}\|_2 
&= \|\sum_{i=1}^k c_i(\mathbf{AX}_{J(S)} - \mathbf{B\hat{X}}_{J(S)})_i\|_2
\leq \varepsilon \sum_{i=1}^k |c_i| \\
&\leq \varepsilon \sqrt{k}  \|\mathbf{c}\|_2 
%\leq \frac{\varepsilon \sqrt{k}}{L(\mathbf{A}X_{J(S)})} \|\mathbf{A}X_{J(S)}\mathbf{c}\|_2 
= \frac{\varepsilon}{L_k(\mathbf{AX}_{J(S)})} \|\mathbf{y}\|_2,
\end{align*}
where the last equality is by definition of $L_k$. It follows that for all $S \in E$,
\begin{align}\label{rhs222}
d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_{S'}\}) 
\leq \frac{\varepsilon}{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) }.
%\leq \frac{\varepsilon}{L(\mathbf{AX}_{J(S)})}.
%&< \left( \frac{ L_2(\mathbf{A}) }{ 2^{|E|} \max_{j \in [m]} \|\mathbf{A}_j\|_2 } \right) \frac{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) } { L(\mathbf{AX}_{J(S)}) } \min_{\substack{F \subseteq E}} r( \{ \text{\rm Span}\{\mathbf{A}_{S}\} \}_{S \in F}) \\
\end{align}
The result then follows by  Lem.~\ref{MainLemma}, with the map $\pi: E \to {[m] \choose k}$ defined by the association $S \mapsto S'$ above.
%\begin{align}
%\delta= \frac{\varepsilon}{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) } = \frac{ C \varepsilon }{ \tilde C } < \frac{L_2(\mathbf{A})}{\tilde C}
%\end{align}
\end{proof}
